---
title: "Brief Introduction to R"
output: rmarkdown::github_document
---

## Installing Libraries in R

```{r, cache = TRUE, echo = TRUE}
#install.packages("dplyr")
#install.packages("RMySQL")
#install.packages("lubridate")
#install.packages("openxlsx")
#install.packages("tidyverse")
#install.packages("stringr")
#install.packages("readr")
```

## Loading Libraries in R

The function `require()` is an equivalent of `library()` only that it outputs a warning and continues if the package is not found.

```{r, cache = TRUE, echo = TRUE, warning = FALSE}
require(dplyr) 
require(RMySQL)
require(lubridate)
require(openxlsx)
#require(tidyverse)
require(stringr)
require(readr)
require(ggplot2)
```

## Help in R - Documentation of every package

```{r, cache = TRUE, echo = TRUE}
?head
```

## Get and set the working directory
```{r, cache = TRUE, echo = TRUE}
getwd()
setwd("~/Documents/")
```


## Data types and structures in R

### Strings - Character

```{r, cache = TRUE, echo = TRUE}
string <- "This is a string"
string
class(string)
length(string)
nchar(string)
```

### Double

```{r, cache = TRUE, echo = TRUE}
number <- 234
number
class(number)
typeof(number)

number <- 234.0
number
class(number)
typeof(number)

number <- 1/8
number
class(number)
typeof(number)
```

### Integers

```{r, cache = TRUE, echo = TRUE}
integer <- 2L
integer
class(integer)
```

### Logical

```{r, cache = TRUE, echo = TRUE}
logical <- FALSE
logical
class(logical)
logical*1
as.logical(1)
```

### Complex

```{r, cache = TRUE, echo = TRUE}
complex <- 1+4i
complex
class(complex)
complex*1
```

### Factor - Label Encoding

```{r, cache = TRUE, echo = TRUE}
factor_1 <- c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun", "Wed", "Thu", "Fri", "Sat", "Sun", "Wed", "Thu", "Wed", "Thu")
factor_1 <- factor(factor_1)
factor_1
levels(factor_1)
class(factor_1)
as.numeric(factor_1)
```

### Ordered Factors - Label Encoding

```{r, cache = TRUE, echo = TRUE}
factor_2 <- c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun", "Wed", "Thu", "Fri", "Sat", "Sun", "Wed", "Thu", "Wed", "Thu")
factor_2 <- ordered(factor_1, levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))
factor_2
class(factor_2)
typeof(factor_2)
as.numeric(factor_2)
```

## Vectors

The data types we saw earlier are **atomic vectors**. A vector is a dynamic array of a unique data type. Is the most common structure in R. Is the essence in R. Most functions in R are vectorized (input: vector) so that we don't have to perform a cycle.

```{r, cache = TRUE, echo = TRUE}
#Sample function
sample(x = 1:100, size = 8, replace = FALSE)
sample(x = 1:5, size = 8, replace = TRUE)

vector("integer", length = 5)

vector_1 <- 1:5
vector_1
vector_1 <- c(1:5)
vector_1
vector_1 <- c(1,2,3,4,5)
vector_1
class(vector_1)

vector_2 <- c(FALSE, TRUE, FALSE)
vector_2
class(vector_2)

vector_3 <- c("a", "b", "c")
vector_3
class(vector_3)

vector_4 <- c("a", 1, "v")
vector_4
class(vector_4)

#adding vectors
c(vector_4, "adding", "more", "elements", "to", "vector_4")

vector_log <- c(1,1,0,0,1,0,1)
as.logical(vector_log)
```

## Lists

Is a collection of many objects/elements/vectors

```{r, cache = TRUE, echo = TRUE}
list_1 <- list(vector_1, vector_2, vector_3)
list_1
list_1[[2]]
names(list_1)
names(list_1) <- c("first", "second", "third")
list_1
list_1$first
```


## Matrix

```{r, cache = TRUE, echo = TRUE}
matrix(data = 1, nrow = 4, ncol = 5)

vector_5 <- c(1:5)
vector_6 <- c(6:10)
vector_7 <- c(11:15)
matrix_1 <- cbind(vector_5, vector_6, vector_7)
matrix_1
class(matrix_1)
```

## Data Frames

```{r, cache = TRUE, echo = TRUE}
df <- data.frame(
  col1 = c("This", "is", "a", "vector", "of", "strings"),
  col2 = 1:6,
  col3 = letters[1:6],
  stringsAsFactors = FALSE
)
df_copy <- data.frame(
  col1 = c("This", "is", "a", "vector", "of", "strings"),
  col2 = 1:6,
  col3 = letters[1:6]
)
df
str(df)
names(df)
names(df) <- c("Column1", "Column2", "Column3")
names(df)
colnames(df)
head(df, 10)
tail(df, 5)
nrow(df)

#Adding more elements to data frame

#New Column
df$Column4 <- 11:16

new_elements <- c("new_string", 19, "z")
df <- rbind(df, new_elements)
df_copy <- rbind(df_copy, new_elements)
df
df_copy

```


### Indexing

```{r, cache = TRUE, echo = TRUE}
len_4 <- length(vector_4)

vector_4
len_4

vector_4[1]
vector_4[2:len_4]
```


```{r, cache = TRUE, echo = TRUE}
df
df[1,1]
df[1,2]
df[1,]
df[,1]
df$col1
df$col1[3]
df[,"Column1"]
```

## Functions of base R

```{r, cache = TRUE, echo = TRUE}
is.na(df_copy)
colSums(is.na(df_copy))
is.na(df_copy$col1)==TRUE
df_copy[is.na(df_copy$col1),]
df_copy[is.na(df_copy$col1)==FALSE,]

#loading diamonds dataset
table(diamonds$cut, diamonds$color)
```


## Defining new functions

```{r, cache = TRUE, echo = TRUE}
#Vectorized funcion - sum and length are vectorized
find_mean <- function(x){
  return(sum(x)/length(x))
}
vec <- c(1:10)
find_mean(vec)


find_sample <- function(x){
  for_index <- sample(1:nrow(x), size = 10, replace = FALSE)
  new_df <- x[for_index,]
  return(new_df)
}
data <- data.frame(
  a = 1:10,
  b = sample(c("GT", "US", "CA"), size = 10, replace = TRUE)
)
find_sample(data)
```

## Apply vs Loop
```{r, cache = TRUE, echo = TRUE}
generate_df <- function(x){
  return(
    data.frame(
      a = sample(letters, size = 10, replace = TRUE),
      b = sample(1:10, size = 10, replace = TRUE)
    )
  )
}
list_2 <- lapply(1:20100, generate_df)

#First, going through a loop
print("Results using a loop")
result_list <- list()
system.time(for(i in 1:length(list_2)){
  result_list[[i]] <- find_sample(list_2[[i]])
})

#Now we are going to use lapply function
print("Results using apply")
system.time(lapply(list_2, find_sample))
```
